### 2020-01-16 [2017-12-15]_HPC_(variance_models)/Presentation/presentation.tex
Created paper repository with all of my work.

% Template for overheads, landscape, fancy VT footlines.
%
\magnification=1800 \rightskip=0pt plus 5em
\hsize=9.0truein \vsize=6.5truein
\parindent=0pt \parskip=0pt plus 4truept
\baselineskip=21truept plus 3truept minus 2truept \lineskiplimit=3truept
\lineskip=3truept plus 3truept \tolerance=1000

%fonts
\font\helvr=phvr at 18truept
\font\helvb=phvb at 18truept
\font\timesi=ptmri at 18truept
\font\helvbXX=phvb at 20truept
\font\helvbXXIV=phvb at 24truept
\font\newfont=pbkli

%% Math blackboard fonts {\bb}
\font\tenamsb=msbm10 \font\sevenamsb=msbm7 \font\fiveamsb=msbm5
\newfam\bbfam
\textfont\bbfam=\tenamsb
\scriptfont\bbfam=\sevenamsb
\scriptscriptfont\bbfam=\fiveamsb

\def\bb{\fam\bbfam}

\def\rm{\fam=0\helvr} \let\bf=\helvb \let\tt=\helvb \let\it=\timesi
\let\sl=\timesi

%definitions
\def\cl#1{\centerline{#1}}
\def\head#1{\centerline{\helvbXXIV#1}}
{\obeyspaces\global\let =\ }
\def\verbatim{\parindent=0pt\tt\obeylines\obeyspaces}
\def\l{\hfill\&}        
\def\shift{\hsize=8.5truein\parindent=25truept\indent}
\def\bull{\par\hangindent=20pt\hangafter=1
\indent\hbox to 20pt{\hfil$\bullet$ }\ignorespaces}

% for spacing in tables
\newdimen\digitwidth \setbox0=\hbox{\rm0} \digitwidth=\wd0

\input epsf % For Postscript figures:{\epsfxsize=  \epsffile{}}
\helvr

%**********************************  Page 1 ******************************
\cl{\helvbXXIV PREDICTIVE MODELING OF I/O CHARACTERISTICS}\smallskip
\cl{\helvbXXIV IN HIGH PERFORMANCE COMPUTING SYSTEMS}
\bigskip
\bigskip
\cl{\bf Thomas Lux$^1$, Layne Watson$^{123}$, Tyler Chang$^1$,}
\cl{\bf Jon Bernard$^1$, Bo Li$^1$, Li Xu$^4$, Godmar Back$^1$,}
\cl{\bf Ali Butt$^1$, Kirk Cameron$^1$, and Yili Hong$^4$}
\bigskip
\bigskip
\cl{Departments of Computer Science$^1$, Mathematics$^2$,}
\cl{Aerospace \& Ocean Engineering$^3$, and Statistics$^4$}
\bigskip
\cl{Virginia Polytechnic Institute and State University}
\cl{Blacksburg, VA 24061-0106  USA} 
\vfil
\vtop to 0pt{\vskip -30truept
\cl{\epsfxsize=1.3in \epsffile{VT_Logo.eps}}\vss}
\footline={\hfil}
\eject

%% Redefine footline for the rest of the presentation

\footline={\fiverm\folio\quad\leaders\hrule
height3truept\hfil\quad \epsfxsize=.8in \epsffile{VT_Logo_CompSci.eps}}

%**********************************  Page 2 ******************************
\head{Problem Setting}
\bigskip \bigskip
When the same computer is used to execute the same program repeatedly, most measurable performance characteristics will vary. This paper will consider the execution of IOzone, a file I/O benchmark for HPC systems.
\cl{\epsfxsize=5in \epsffile{Raw_Throughput.eps}}
%% \cl{\hbox{\epsfxsize=30truepc\epsffile{triangleplane.eps}}}
This variance in performance can make it difficult to identify an optimal configuration that meets minimum requirements.
\vfil
\eject

%**********************************  Page 3 ******************************
\head{Proposed Approach}
\bigskip \bigskip
Use multivariate interpolation and regression to model the performance of a computer with respect to the system and application configuration.
%% \bull Bulletted text..
\bigskip\bigskip
Multivariate {\it interpolation} and {\it regression} are defined when there exists $f:{\bb R}^d \rightarrow {\bb R}$ and a set $X$ of $n$ points in ${\bb R}^d$ along with response values $f(x)$ for all $x \in X$.
\bigskip\bigskip
{\bf Interpolation} constructs an approximation $\hat f: {\bb R}^d \rightarrow {\bb R}$ such that $\hat f(x) = f(x)$ for all $x \in X$. The form of the true underlying function $f$ is often unknown, however it is still desirable to construct an approximation $\hat f$ with minimum approximation error at $y \notin X$.
\bigskip\bigskip
{\bf Regression} relaxes the conditions of interpolation by minimizing the error in $\hat f$ at $x \in X$ while maintaining some parametric form with parameters $P$. This can be written as $\min_{P} ||\hat f(X) - f(X)||$, where $f(X)$ is a vector of $f(x)$ for all $x \in X$ and $||\cdot||$ is an appropriate measure.
\vfil
\eject

%**********************************  Page 4 ******************************
\head{Regression Algorithms}
\bigskip
{\bf Multivariate Adaptive Regression Splines} (MARS)
is an iterative approximation of the form
$$ B_{2s-1}(x) = B_l(x) [c(x_i-v)]_+ ,$$
$$ B_{2s}(x) = B_k(x) [c(x_i-v)]_- ,$$
where $s$ is the iteration number, $B_l(x)$ and $B_k(x)$ are basis
functions from the previous iteration, $c, v \in {\bb R}$, 
$$w_+ = \cases{ w, & $w \geq 0$ \cr 0, & $w < 0$},$$
and $w_- = (-w)_+$. After iteratively constructing a model, MARS then
iteratively removes basis functions that do not contribute to goodness
of fit.
\medskip
In effect, MARS creates a locally component-wise tensor
product approximation of the data. The computational
complexity of MARS is ${\cal O}(n d m^3)$ where $m$ is the
maximum number of underlying basis functions.
\vfil\eject

%**********************************  Page 4 ******************************
\head{Regression Algorithms}
\bigskip
{\bf Multilayer Perceptron Regressor} (MLP Regressor)
$$ l(u) = \big( u^t W_l \big)_+ ,$$
where $W_l$ is the $i$ by $j$ weight matrix for layer $l$.
\medskip
The multilayer perceptron (MLP) produces a piecewise linear model of the input data. The computational complexity is ${\cal O}(n d m)$, where $m$ is
determined by the sizes of the layers of the network and the stopping
criterion of the error minimizer.
\bigskip
{\bf Support Vector Regressor} (SVR)
$$ p(x)  = \sum_{i=1}^{n}a_i K(x,x^{(i)}) + b ,$$
where $K$ is the selected kernel function, $a \in {\bb R}^n$, $b \in {\bb R}$ are coefficients to be solved for simultaneously. The computational complexity of the SVR is ${\cal O}(n^2dm)$, with $m$ being determined by the minimization convergence criterion.
\vfil\eject

%**********************************  Page 5 ******************************
\head{Interpolation Algorithms}
\bigskip
{\bf Delaunay} constructs a simplicial mesh such that for a $d$-simplex S with vertices $v^{(0)}$, $v^{(1)}$, $\ldots$, $v^{(d)}$, $x \in S$, and data values $f(v^{(i)})$, $i=0$, $\ldots$, $d$, $x$ is a unique convex combination of the vertices, and the interpolant is given by
$$ p(x) = \sum_{i=0}^{d} w_i f(v^{(i)}), $$
where $w_i$ are the convex weights. The computational complexity of the Delaunay triangulation (for the implementation used here) is ${\cal O}(n^{\lceil d/2 \rceil})$.
\bigskip
{\bf Linear Shepard} (LSHEP) blends local linear interpolants and has the form
$$ p(x) = {\sum_{k=1}^{n}W_k(x)P_k(x) \over \sum_{k=1}^{n}W_k(x)} ,$$
where $W_k(x)$ is a locally supported weighting function and $P_k(x)$ is a local linear approximation to the data satisfying $P_k\big(x^{(k)}\big) = f\big(x^{(x)}\big)$. The computational complexity of LSHEP is ${\cal O}(n^2d^3)$.
\vfil \eject

%**********************************  Page 6 ******************************
\head{IOzone Data}
\bigskip \bigskip
This experiment attempts to model {\it throughput variance} as
a function of application and system {\it parameters}.
\medskip
To do so, the {\it IOZone} benchmark is used to read files of varying sizes
on a homogeneous system. 
The variance in the throughput for each read is modelled as a function of
several parameters.
\medskip
The parameters chosen and the values used for them are in the table below:
\smallskip
\offinterlineskip \tabskip=0pt
\def\filler{height2pt&\omit&&\omit&\cr}
\def\tablerule{\noalign{\hrule}}
\centerline{\vbox{
\halign{&\vrule#&\strut\hskip 4pt\hfil#\hskip 4pt\cr
\tablerule\filler
& Parameters \hfill&& Values \hfill&\cr
\filler\tablerule\filler
&\omit\hskip 4pt\hbox{file size being read (in KB)}
&&64, 256, 1024\hfill&\cr
&\omit\hskip 4pt\hbox{record size of file system (in KB)}
&&32, 128, 512\hfill&\cr
&\omit\hskip 4pt\hbox{number of threads for IOZone reader}
&&1, 2, 4, 8, 16, 32, 64, 128, 256\hfill&\cr
&\omit\hskip 4pt\hbox{CPU frequency (in GHz)}
&&1.2, 1.4, 1.5, 1.6, 1.8, 1.9, 2.0, 2.1, 2.3, &\cr
&\hfill
&& 2.4, 2.5, 2.7, 2.8, 2.9, 3.0, 3.001 \hfil&\cr
\filler\tablerule}}}
\medskip
For each combination of the above parameters, 40 runs of IOzone were done and 
the observed throughput variance was computed using:
$$ \sigma^2 = \bigg(\sum_{i=1}^{40}(t_i - \mu)^2 \bigg) / 39 $$
where $t_i$ is the $i$th observed throughput and $\mu$ is the observed
mean over all 40 runs.
\vfil\eject

%**********************************  Page 7 ******************************
\head{Analyzing Predictability: Multi Dimensional Analysis}
\bigskip\bigskip

\item{1.} For all $k = 1$, $\ldots$, $d$ and for all nonempty subsets $F \subset \{1, 2, \ldots, d\}$, reduce the input data to points $(z, f_F(z))$ with $z \in {\bb R}^k$ and $f_F(z) = E\bigl[ \bigl\{ f\bigl(x^{(i)}\bigr) \bigm| \bigl(x^{(i)}_F = z\bigr) \bigr\} \bigr]$, where $E[\cdot]$ denotes the mean and $x^{(i)}_F$ is the subvector of $x^{(i)}$ indexed by $F$.
\bigskip
\item{2.} For all $r$ in $\{5, 10, \ldots, 95\}$, generate $N$ random splits $(train, test)$ of the reduced data with $r$ percentage for training and $100 - r$ percentage for testing.
\bigskip
\item{3.} When generating each of $N$ random $(train, test)$ splits, ensure that all points from $test$ are in the convex hull of points in $train$ (to prevent extrapolation); also ensure that the points in $train$ are well spaced.
\bigskip\bigskip

In order to ensure that training points are well spaced, a statistical method for picking points is used
\bigskip
\itemitem{1.} Generate a sequence of all pairs of points $\bigl(z^{(i_1)},z^{(j_1)}\bigr), \bigl(z^{(i_2)},z^{(j_2)}\bigr), \ldots$ sorted by ascending pairwise Euclidean distance between points, so that $\bigl|\bigl|z^{(i_k)}-z^{(j_k)}\bigr|\bigr|_2 \leq \bigl|\bigl|z^{(i_{k+1})}-z^{(j_{k+1})}\bigr|\bigr|_2$.
\bigskip
\itemitem{2.} Sequentially remove points from candidacy until only $|train|$ remain by randomly selecting one point from the pair $\bigl(z^{(i_m)}, z^{(j_m)}\bigr)$ for $m = 1,\ldots$ if both $z^{(i_m)}$ and $z^{(j_m)}$ are still candidates for removal.

\vfil\eject

%**********************************  Page 8 ******************************
\head{Increasing Samples: Results for Predicting Throughput Mean}
\bigskip\bigskip\bigskip
\cl{\epsfxsize=5in \epsffile{Mean_TT_Ratio.eps}}
\vfil\eject

\head{Increasing Samples: Results for Predicting Throughput Variance}
\bigskip\bigskip\bigskip
\cl{\epsfxsize=5in \epsffile{Var_TT_Ratio.eps}}
\vfil\eject

%**********************************  Page 10 ******************************
\head{Increasing Dimension: Results for Predicting Throughput Mean}
\bigskip\bigskip\bigskip
\cl{\epsfxsize=5in \epsffile{Mean_Dim.eps}}
\vfil\eject

\head{Increasing Dimension: Results for Predicting Throughput Variance}
\bigskip\bigskip\bigskip
\cl{\epsfxsize=5in \epsffile{Var_Dim.eps}}
\vfil\eject

%**********************************  Page 11 ******************************
\head{Conclusion and Future Work}
\bigskip\bigskip\bigskip
Multivariate models of HPC system performance can effectively predict
I/O throughput mean and variance. These multivariate techniques
significantly expand the scope and portability of statistical models
for predicting computer system performance over previous work.
\bigskip\bigskip\bigskip
Delaunay model can make predictions for 821 system configurations with less than 5\% error when trained on only 43 configurations.
\bigskip\bigskip\bigskip
Multivariate methods should be applied to predict distributions instead of summary statistics. This would allow for much deeper insight.
\vfil\eject

%**********************************  Page 12 ******************************
\head{Acknowledgements}
\bigskip \bigskip
The data shown here was collected for the VarSys project at Virginia Tech.
\vfil\eject

\bye