### 2020-01-16 [2018-01-09]_ACMSE_(box_meshes)/paper_header.tex

%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
\documentclass[sigconf]{acmart}
%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.

\usepackage{booktabs} % For formal tables
\usepackage{tikz}     % For drawing example boxes
\usepackage{graphicx}
\usetikzlibrary{positioning}

\begin{document}

%%%% Proceedings format for SIGPLAN conferences 
% \documentclass[sigplan, anonymous, review]{acmart}
%%%% Proceedings format for SIGCHI conferences
% \documentclass[sigchi, review]{acmart}
%%%% To use the SIGCHI extended abstract template, please visit
% https://www.overleaf.com/read/zzzfqvkmrfzn
% Copyright
%% \setcopyright{none}
% \setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}
% DOI
% \acmDOI{}
% ISBN
% \acmISBN{}
%Conference
% \acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA} % \acmConference[ACMSE 2018]{ACM Southeastern conference}{March 2018}{Richmond, Kentucky USA} 
% \acmYear{2018}
% \copyrightyear{2018}
% \acmArticle{4}
% \acmPrice{0.00}
% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}

\copyrightyear{2018} 
\acmYear{2018} 
\setcopyright{acmlicensed}
\acmConference[ACM SE '18]{ACM SE '18: Southeast Conference}{March 29--31, 2018}{Richmond, KY, USA}
\acmBooktitle{ACM SE '18: ACM SE '18: Southeast Conference, March 29--31, 2018, Richmond, KY, USA}
\acmPrice{15.00}
\acmDOI{10.1145/3190645.3190687}
\acmISBN{978-1-4503-5696-1/18/03}

\title{Novel Meshes for Multivariate Interpolation and Approximation}
%\titlenote{Produces the permission block, and copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as \texttt{acmart.pdf} document}

\author{Thomas C. H. Lux}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \department{Dept. of Computer Science}
  \institution{Virginia Polytechnic Institute and State University}
  \city{Blacksburg} 
  \state{Virginia} 
  \postcode{24060}
}
\email{tchlux@vt.edu}

\author{Layne T. Watson}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \department{Depts. of Computer Science, Mathematics, and Aerospace \& Ocean Engineering}
  \institution{Virginia Polytechnic Institute and State University}
}

\author{Tyler H. Chang}
\author{Jon Bernard}
\author{Bo Li}
\author{Xiaodong Yu}
\affiliation{%
  \department{Dept. of Computer Science}
  \institution{Virginia Polytechnic Institute and State University}
}

\author{Li Xu}
\affiliation{%
  \department{Dept. of Statistics}
  \institution{Virginia Polytechnic Institute and State University}
}

\author{Godmar Back}
\author{Ali R. Butt}
\author{Kirk W. Cameron}
\author{Danfeng Yao}
\affiliation{%
  \department{Dept. of Computer Science}
  \institution{Virginia Polytechnic Institute and State University}
}

\author{Yili Hong}
\affiliation{%
  \department{Dept. of Statistics}
  \institution{Virginia Polytechnic Institute and State University}
}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{T. Lux et al.}

\begin{abstract}

A rapid increase in the quantity of data available is allowing all fields of science to generate more accurate models of multivariate phenomena. Regression and interpolation become challenging when the dimension of data is large, especially while maintaining tractable computational complexity. This paper proposes three novel techniques for multivariate interpolation and regression that each have polynomial complexity with respect to number of instances (points) and number of attributes (dimension). Initial results suggest that these techniques are capable of effectively modeling multivariate phenomena while maintaining flexibility in different application domains.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
% \begin{CCSXML}
% <ccs2012>
%  <concept>
%   <concept_id>10010520.10010553.10010562</concept_id>
%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%   <concept_significance>500</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010575.10010755</concept_id>
%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
%   <concept_significance>300</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010553.10010554</concept_id>
%   <concept_desc>Computer systems organization~Robotics</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10003033.10003083.10003095</concept_id>
%   <concept_desc>Networks~Network reliability</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
% </ccs2012>  
% \end{CCSXML}

% \ccsdesc[500]{Computer systems organization~Embedded systems}
% \ccsdesc[300]{Computer systems organization~Redundancy}
% \ccsdesc{Computer systems organization~Robotics}
% \ccsdesc[100]{Networks~Network reliability}

\keywords{Interpolation, Approximation, Splines, Multivariate, Regression}

\maketitle