### 2020-03-30 [2019-05]_QIP_(quantum_ls)/7-conclusion.tex

\section{Conclusion and Future Work}

In this work, a methodology is proposed for encoding polynomial SOS minimization for both fixed-point decimal and integer valued problems.
To achieve this, the squared error function is mapped to the QUBO model, which is isomorphic to the Ising-Hamiltonian model, using a two's complement encoding and the quadratization scheme of Rosenberg \cite{rosenberg1975reduction}.
The current state-of-the-art D-Wave Chimera architecture is able to solve the proposed embedding for a wide variety of problems, some of which do not admit analytic classical solutions, with the primary limitation being the qubit requirement for large systems.
The fact that the QA ground state is a minimum forward error solution carries tremendous advantages for ill-conditioned problems.
In general, these results accompany and extend the recently published results of Borle et al. \cite{borle2019analyzing} and Chang et al. \cite{chang2019quantum}.

Overall, this is a powerful methodology with tremendous applications in the areas of numerical analysis, machine learning, computer security, and general scientific computing.
As the state-of-the-art in quantum annealing hardware continues to evolve, this methodology should scale to increasingly complex problems.