### 2020-03-30 [2019-05]_QIP_(quantum_ls)/fig1.tex

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{fig1a.pdf}$\qquad$
\includegraphics[width=0.35\textwidth]{fig1b.pdf}
\caption{Finding a physical embedding: this figure shows a toy 32 qubit system demonstrating the Chimera graph structure (left) and a synthesized dense network on an individual unit cell (right). The Chimera graph consists of bipartite 8-qubit unit cells as shown, with sparse connections between cells.
To synthesize a dense 4-qubit network on any individual unit cell, negative coupler weights (shown in bold) are added.
I.e., $J_{0,4} = J_{1,5} = J_{2,6} = J_{3,7} < 0$.
}
\label{fig:chains}
\end{figure}