### 2022-03-09 [2020-08]_ACMTOMS_(MQSI)/paper/mqsi20.tex

% Algorithm XXXX: MQSI---Monotone quintic spline interpolation,
% Thomas C.H. Lux, Layne T. Watson, Tyler H. Chang,
% and William Thacker
% ACM Trans. Math. Software, ?/?/202?.
% 
%
\magnification=1000 \hsize=30pc \vsize=47pc
\parindent=10pt \parskip=0pt plus 3pt
\baselineskip=13pt plus 2pt minus 1pt \lineskiplimit=2pt
\lineskip=2pt plus 2pt %\raggedright
\tolerance=800
%\abovedisplayskip=6pt plus 6pt minus 3pt
%\abovedisplayshortskip=0pt plus 3pt
%\belowdisplayskip=6pt plus 6pt minus 3pt
%\belowdisplayshortskip=6pt plus 2pt minus 3pt

% \cite is a placeholder to use for references while writing.
\def\cite#1{[{\tt REF}]}

% Color definitions.
\input colordvi
\def\BEGINC{\textPeach} \def\ENDC{\textBlack}
\def\red#1{\BEGINC#1\ENDC}
\let\beginred=\BEGINC
\let\endred=\ENDC

% Commands for inserting an 'algorithm' section.
\def\thickrule{\hrule\hrule\hrule}
\def\codent{\hskip 1.0em} % <- 'tab' for indentation.

% fonts
\font\itVIII=cmti8 \font\slVIII=cmsl8 \font\bfVIII=cmbx8
\font\ttVIII=cmtt8 \font\caps=cmcsc10 \font\capsVIII=cmcsc10 at 8pt
\font\ttXVIII=cmtt12 at 18pt \font\eightmi=cmmi8 \font\eightsy=cmsy8
\font\rmeight=cmr8  \def\rmVIII{\rmeight \baselineskip=10pt plus 2pt
minus 1pt \lineskiplimit=2pt \lineskip=2pt plus 1pt
\parskip=0pt \textfont0=\rmeight \textfont1=\eightmi \textfont2=\eightsy}
\font\hlvXVIII=phvr at 18pt \font\hlvVIII=phvr at 8pt \font\hlv=phvr

% math blackboard font
\font\tenamsb=msbm10 \font\sevenamsb=msbm7 \font\fiveamsb=msbm5
\newfam\bbfam
\textfont\bbfam=\tenamsb
\scriptfont\bbfam=\sevenamsb
\scriptscriptfont\bbfam=\fiveamsb
\def\bbb{\fam\bbfam}

%Postscript figures.
\input epsf
\def\lcaption#1#2{{\noindent\rmVIII Fig.~#1. #2\par}}
\def\ltabcaption#1#2{{\noindent\rmVIII Table #1. #2\par}}

% other definitions.
\def\h{\par\hangindent=10pt\hangafter=1\noindent}
\def\heading#1{\par\bigskip\leftline{\hlv#1}\nobreak\smallskip
\noindent\ignorespaces}
\newdimen\headdigit{\setbox11=\hbox{\hlv 8. }\global\headdigit=\wd11}
\def\footnoterule{\bigskip\kern-3pt \hrule \kern 2.6pt \smallskip}
\def\bullitem{\par\vskip 2pt plus 2pt\hangindent=15pt \hangafter=1
\noindent\hbox to 15pt{---\hfil}\ignorespaces}
\def\bulllitem{\par\vskip 2pt plus 2pt\hangindent=30pt
\hangafter=1 \noindent\hskip 15pt\hbox to
15pt{---\hfil}\ignorespaces}
\def\Th#1 #2\par{\par\medskip{\it{\caps#1}. #2}\par\medskip}
\def\btable#1{\topinsert \rmVIII \let\bf=\bfVIII \textfont0=\scriptfont0
\scriptfont0=\scriptscriptfont0 \textfont1=\scriptfont1
\scriptfont1=\scriptscriptfont1 \centerline{#1}}
\let\xpar=\par
\def\ds{\displaystyle}\def\ab{\allowbreak}

% math definitions
\def\tn#1{\left\Vert #1\right\Vert_2}
\def\mn#1{\left\Vert #1\right\Vert_\infty}
\def\lee{\mathrel{\vcenter{\hbox{$\scriptstyle\mathord<$}\nointerlineskip
\vskip 1pt\hbox{$\scriptstyle\mathord=$}}}}
\def\gee{\mathrel{\vcenter{\hbox{$\scriptstyle\mathord>$}\nointerlineskip
\vskip 1pt\hbox{$\scriptstyle\mathord=$}}}}
\def\real{\mathop{\rm I\!R}\nolimits}
\def\R{{\bf R}}
\def\Rd{{\bf R}^d}
\def\Rdd{{\bf R}^{d\times d}}
%\def\bull{\hfill \vrule height6pt width6pt depth0pt}
\def\:{\mathrel{:\mathord=}}

% OUTPUT
\def\rightheadline{\hfill\hlvVIII\rightrh\hskip20pt$\scriptstyle\bullet$\hskip
20pt\folio}
\def\leftheadline{\hlvVIII\folio\hskip20pt$\scriptstyle\bullet$\hskip
20pt\leftrh\hfill}
\footline={\hfill}%no page number on first page
\output{\plainoutput}
\def\plainoutput{\shipout\vbox{\makeheadline\pagebody\makefootline}
  \advancepageno \global\footline={\hfill}
  \ifodd\pageno
        \global\headline={\rightheadline}
  \else
        \global\headline={\leftheadline}
  \fi
  \ifnum\outputpenalty>-20000 \else\dosupereject\fi}
\def\pagebody{\vbox to\vsize{\boxmaxdepth\maxdepth \pagecontents}}
\def\makeheadline{\vbox to 0pt{\vskip-22.5pt
  \line{\vbox to8.5pt{}\the\headline}\vss}\nointerlineskip}
\def\makefootline{\baselineskip24pt\vskip-6pt\line{\the\footline}}
\def\dosupereject{\ifnum\insertpenalties>0pt
  \line{}\kern-\topskip\nobreak\vfill\supereject\fi}
%\def\footnoterule{\vskip10pt\kern-3pt \hrule width 3em \kern 2.6pt}

\def\leftrh{Lux et al.}
\def\rightrh{Algorithm XXXX: MQSI}

{\baselineskip=24pt
\leftline{\hlvXVIII Algorithm XXXX: MQSI---Monotone Quintic}
\leftline{\hlvXVIII Spline Interpolation}}
\bigskip\medskip
\leftline{\hlv THOMAS C. H. LUX and LAYNE T. WATSON}
\smallskip
\leftline{\hlv Virginia Polytechnic Institute and State University}
\smallskip
\leftline{\hlv TYLER H. CHANG}
\smallskip
\leftline{\hlv Argonne National Laboratory}
\smallskip
\leftline{\hlv WILLIAM I. THACKER}
\smallskip
\leftline{\hlv Winthrop University}
\bigskip

\hrule\bigskip\smallskip \footnote{}{\hskip
-\parindent{\parindent=0pt\rmVIII
This work was supported by National Science Foundation Grants CNS-1565314,
CNS-1838271, and DGE-154362.
Authors' addresses:
T. C. H. Lux, L. T. Watson, Departments of Computer Science, Mathematics,
and Aerospace and Ocean Engineering, Virginia Polytechnic Institute \&
State University, Blacksburg, VA 24061; e-mails: {\ttVIII tchlux@vt.edu,
ltwatson@computer.org}; T. H. Chang, Mathematics and Computer Science
Division, Argonne National Laboratory, 9700 South Cass Avenue, Bldg. 240,
Lemont, IL 60439; e-mail: {\ttVIII thchang@vt.edu}; W. I. Thacker, Winthrop
University, Rock Hill, SC 29733; {\ttVIII thackerw@winthrop.edu}.

\xpar
% Permission to make digital/hard copy of
%part or all of this work for personal or classroom use is granted
%without fee provided that the copies are not made or distributed for
%profit or commercial advantage, the copyright notice, the title of
%the publication, and its date appear, and notice is given that
%copying is by permission of the ACM, Inc. To copy otherwise, to
%republish, to post on servers, or to redistribute to lists, requires
%specific permission and/or fee. \xpar\copyright\ 2018 by the
%Association for Computing Machinery, Inc. \xpar 
}}

% Abstract and Introduction
{\rmVIII\parindent=0pt MQSI is a Fortran 2003 subroutine for constructing
monotone quintic spline interpolants to univariate monotone data. Using sharp
theoretical monotonicity constraints, first and second derivative estimates
at data provided by a quadratic facet model are refined to produce a
univariate $\scriptstyle C^2$ monotone interpolant.  Algorithm and implementation
details, complexity and sensitivity analyses, usage information, a
brief performance study, and comparisons with other spline approaches are included.

\medskip
Categories and Subject Descriptors: G.1.1 [{\bfVIII Numerical Analysis}]:
Interpolation --- Spline and piecewise polynomial interpolation;
J.2 [{\bfVIII Computer Applications}]: Physical Science and Engineering
--- {\itVIII Mathematics};
G.4 [{\bfVIII Mathematics of Computing}]: Mathematical Software

%% 10002950.10003705
%% Mathematics of computing~Mathematical software
%% 
%% 10002950.10003648.10003649.10003657.10003659
%% Mathematics of computing~Spline models
%% 
%% 10002950.10003714.10003715.10003722
%% Mathematics of computing~Interpolation
%% 
%% 10002950.10003714.10003715
%% Mathematics of computing~Numerical analysis
%% 
%% 10002950.10003648.10003649.10003657
%% Mathematics of computing~Nonparametric representations

\medskip
General Terms: Algorithms, Monotonicity, Spline

\medskip
Additional Key Words and Phrases: quintic spline, interpolation

}

\bigskip\hrule\bigskip

%% Section 1
\input 1-introduction

%% Section 2
\input 2-monotone-quintic

%% Section 3
\input 3-spline-rep

%% Section 4
\input 4-analysis

%% Section 5
\input 5-applications

%% Acknowledgements
%% \heading{Acknowledgments}
%% This work was supported by the National Science Foundation Grants CNS-1565314 and CNS-1838271.

% ====================================================================
%                            Bibliography
% \vfil\goodbreak
\heading{\hlvVIII BIBLIOGRAPHY} {
\baselineskip=2pt%\parskip=100mm
\frenchspacing
% Temporary 'caps', 'sl', and 'rm'.
\let\caps=\capsVIII
\let\sl=\slVIII
\let\rm=\rmeight \rmVIII

% \ys makes a half-em horizontal space.
\def\ys{\hskip 1em minus .5em}

%\refj is for regular journal articles
%  #1 -- Author(s)
%  #2 -- Year
%  #3 -- Title
%  #4 -- Journal
%  #5 -- Volume, number, and pages
\def\refj#1#2#3#4#5{\noindent\hangindent=10pt\hangafter=1
{\caps #1}\ys #2.\ys {\rm #3}. {\sl #4}, #5.\vskip 6pt minus 3pt}

%\refb is for books
%  #1 -- Author(s)
%  #2 -- Year
%  #3 -- Title
%  #4 -- Publisher, City, State
\def\refb#1#2#3#4{\noindent\hangindent=10pt\hangafter=1
{\caps #1}\ys #2.\ys{\sl #3}. {\rm #4}.\vskip 6pt minus 3pt}

%\reft is for tech. reports, theses, and conferences
%  #1 -- Author(s)
%  #2 -- Year
%  #3 -- Title
%  #4 -- Technical report ID, Publisher, City, State
\def\reft#1#2#3#4{\noindent\hangindent=10pt\hangafter=1
{\caps #1}\ys #2.\ys{\rm #3}. {\rm #4}.\vskip 6pt minus 3pt}

%% Proves approximation error bounds for monotone spline interpolation of data from monotone functions.
\refj{DeVore, R. A.}{1977}{Monotone approximation by splines}{SIAM Journal on Mathematical Analysis}{8(5):891--905}

%% Accompanies the previous error bounds in the context of monotone polynomial interpolation of data from a monotone function.
\refj{DeVore, R. A.}{1977}{Monotone approximation by polynomials}{SIAM Journal on Mathematical Analysis}{8(5):906--921}

%% Develops a quadratic spline curve fitting algorithm that maintains local convexity while also proving some error bounds. This method avoids inserting additional knots, but does not interpolate the data.
\refj{Beatson, R. K.}{1982}{Monotone and convex approximation by splines: error estimates and a curve fitting algorithm}{SIAM Journal on Numerical Analysis}{19(6):1278--1285}

%% Proves necessary and sufficient conditions for a closed form solution to monotone C1 quadratic rational spline interpolation. Does not provide formal algorithm, nor code.
\refj{Gregory , J. A.,Delbourgo,  R.}{1982}{Piecewise rational quadratic interpolation to monotonic data}{IMA Journal of Numerical Analysis}{2(2):123--130}

%% Iterates towards a solution of nonlinear equations that ensure C2 continuity of quadratic rational splines. The solution they iterate towards exists and is unique. They provide proofs and methods for setting derivatives at the ends, no formal algorithms or code.
\refj{Delbourgo, R., Gregory, J. A.}{1983}{$C^2$ rational quadratic spline interpolation to monotonic data}{IMA Journal of Numerical Analysis}{3(2):141--152}

%% Proves existence of splines that maintain monotonicity and convexity while having fixed continuity and order.
\refj{Costantini, P.}{1986}{On monotone and convex spline interpolation}{Mathematics of Computation}{46(173):203--214}

%% Proposes an algorithm for constructing C2 Hermite monotone spline interpolants using proven sufficient, but not necessary conditions. As these bounds are not sharp, the space of allowable derivative and second derivative values at knots is unnecessarily shrunk. An algorithm is provided, but no code. Recommends against the direct optimization for smoothness parameters like L2, instead promoting problem specific definitions of 'geometric niceness'.
\refj{Dougherty, R. L., Edelman, A., Hyman, J. M.}{1989}{Nonnegativity-, monotonicity- convexity-preserving cubic and quintic Hermite interpolation}{Mathematics of Computation}{52(186):471--494}

%% Proves a simple method for determining the existence of a monotone C2 cubic interpolating spline, but functions of this kind do not always exist for arbitrary convex monotone data sets.
\refj{Fiorot, J. C., Tabka, J.}{1991}{Shape-preserving $C^2$ cubic polynomial interpolating splines}{Mathematics of Computation}{57(195):291--298}

%% Continuous previous work by incorporating a cubic numerator, proves shape preservation when the correct value for a tension parameter is set but no method for setting that parameter. Accuracy depends on accurate derivative evaluations. No code nor formal algorithm is provided.
\refj{Delbourgo, R.}{1993}{Accurate $C^2$ rational interpolations in tension}{SIAM Journal on Numerical Analysis}{30(2):595--607}

%% Presents several nonlinear boundary conditions based on the median function to establish a fourth order monotonicity preserving constraints relying on divided difference initial derivative estimates. Overall generates a monotone C1 cubic interpolant to data.
\refj{Huynh, H. T.}{1993}{Accurate monotone cubic interpolation}{SIAM Journal of Numerical Analysis}{30(1):57--100}

%% Proposes a cubic C2 shape preserving spline method with free parameters (mentions quintic splines are necessary for globally C2 monotone splines). Sufficient (but not necessary) conditions are provided for monotonicity.
\refj{Pruess, S.}{1993}{Shape preserving $C^2$ cubic spline interpolation}{IMA Journal of Numerical Analysis}{13(4):493--507}

%% Constructs a C2 cubic spline with two additional knots per interval. No code (paywall).
\refj{Manni, C., Sablonni\`ere, P.}{1997}{Monotone interpolation of order 3 by $C^2$ cubic splines}{IMA Journal of Numerical Analysis}{17(2):305--320}

%% Uses sufficient conditions to generate C2 cubic spline. No code (paywall).
\refj{Manni, C.}{2001}{On shape preserving $C^2$ Hermite interpolation}{BIT Numerical Mathematics}{41(1):127--148}

%% Constructs C3 interpolating splines by progressively increasing tension through Bezier control polygons until shape parameters are satisfied, no code is provided.
\refj{Cravero, I., C. Manni}{2003}{Shape-preserving interpolants with high smoothness}{Journal of computational and applied mathematics}{157(2):383--405}

%% No extra knots, rational functions, sufficiency and necessity proven, works for unequally spaced data, no code associated with the paper, no algorithm provided for the initialization for free shape parameters.
\refj{Abbas M., Majid, A. A. and Ali, J. M.}{2012}{Monotonicity-preserving $C^2$ rational cubic spline for monotone data}{Applied Mathematics and Computation}{219(6):2885--2895}

%% Use sufficient conditions for monotonicity and convexity to form a globally weighted combination of C1 quadratic splines that satisfy shape constraints.
\refj{Kvasov, B. I.}{2014}{Monotone and convex interpolation by weighted quadratic splines}{Advances in Computational Mathematics}{40(1):91--116}

%% Uses sufficient conditions over three free parameters to construct C2 rational quartic shape preserving splines.
\refj{Zhu, Y.,Han, X.}{2015}{$C^2$ rational quartic interpolation spline with local shape preserving property}{Applied Mathematics Letters}{46:57--63}

%% Uses sufficient conditions over two free parameters to construct C2 rational quartic shape preserving splines.
\refj{Zhu, Y.,Han, X.}{2015}{Shape preserving $C^2$ rational quartic interpolation spline with two parameters}{International Journal of Computer Mathematics}{92(10):2160--2177}

%% Proposes a method for making arbitrarily smooth shape preserving spline interpolants, but uses a centered quadratic to estimate derivatives and a sufficient set of bounding conditions to ensure shape preservation in practice. No necessary conditions are applied, no algorithm is provided, neither is any code provided.
\refj{Han, X.}{2018}{Shape-preserving piecewise rational interpolation with higher order continuity}{Applied Mathematics and Computation}{337:1--13}

%% -------------------------------------------------------------------

% berglund2009planning
\refj{Berglund, T., Brodnik, A., Jonsson, H., Staffanson, M., and
  Soderkvist, I.}{2009}{Planning smooth and obstacle-avoiding B-spline
  paths for autonomous mining vehicles}{IEEE Transactions on
  Automation Science and Engineering}{7(1):167--172}

% brennan2019measure
\refj{Brennan, A.}{2020}{Measure, modulation and metadesign: NC
  fabrication in industrial design and architecture}{Journal of Design
  History}{33(1):66--82}

% deboor1978practical
\refb{de Boor, C.}{1978}{A Practical Guide to Splines}{Springer, Verlag,
New York}

% cameron2019moana
\refj{Cameron, K. W., Anwar, A., Cheng, Y., Xu, L., Li, B., Ananth, U.,
Bernard, J., Jearls, C., Lux,  T. C. H., Hong, Y., Watson, L. T., and Butt,
A. R.}{2019}{MOANA: modeling and analyzing I/O variability in parallel
system experimental design}{IEEE Transactions on Parallel and Distributed
Systems}{30(8):1843--1856}

% fritsch1980monotone
\refj{Fritsch., F and Carlson., R.}{1980}{Monotone piecewise cubic
  interpolation}{SIAM Journal on Numerical Analysis}{17(2):238--246}

% fritsch1982LLNL
\reft{Fritsch, F. N.}{1982}{Piecewise cubic Hermite interpolation
  package, final specifications}{Computer Documentation UCID-30194,
  Lawrence Livermore National Laboratory}

% gregory1985shape
\reft{Gregory, J. A.}{1985}{Shape preserving spline
  interpolation}{19850020252, Brunel University}

% herman2015techniques
\reft{Herman, D. L., and Oftedal, M. J.}{2015}{Techniques and
  workflows for computer graphics animation systems}{US Patent
  9,216,351}

% hess1994positive
\refj{He{\ss}, W. and Schmidt, J. W.}{1994}{Positive quartic, monotone
  quintic $\scriptstyle C^2$-spline interpolation in one and two
  dimensions}{Journal of Computational and Applied Mathematics}{55(1):51--67}

% knott2012interpolating
\reft{Knott, G. D.}{2012}{Interpolating cubic splines}{Springer Science
  \& Business Media, Volume 18}

% leitenstorfer2006generalized
\refj{Leitenstorfer, F. and Tutz, G.}{2006}{Generalized monotonic
  regression based on B-splines with an application to air pollution
  data}{Biostatistics}{8(3):654--673}

% lux2020thesis
\reft{Lux, T. C. H.}{2020}{Interpolants and error bounds for modeling
  and predicting variability in computer systems}{Thesis, Virginia
  Polytechnic Institute and State University, Blacksburg, Virginia}

% mcallister1981algorithm
\refj{McAllister, D. F. and Roulier, J. A.}{1981}{An algorithm for 
  computing a shape-preserving osculatory quadratic spline}{ACM 
  Transactions on Mathematical Software (TOMS)}{7(3):331--347}

% mcallister1981shape
\refj{McAllister, D. F. and Roulier, J. A.}{1981}{Algorithm 574:
  Shape-Preserving Osculatory Quadratic Splines [E1, E2]}{ACM
  Transactions on Mathematical Software (TOMS)}{7(3):384--386}

% moler2004numerical
\refb{Moler, C. B.}{2008}{Numerical Computing with {\caps Matlab}}{
  The MathWorks Inc., Natick, Massachusetts}

% piah2011improved
\refj{Piah, A. R. M. and Unsworth, K.}{2011}{Improved sufficient
  conditions for monotonic piecewise rational quartic
  interpolation}{Sains Malaysiana}{40(10):1173--1178}

% quint2003scalable
\refj{Quint, A.}{2003}{Scalable vector graphics}{IEEE
  MultiMedia}{10(3):99--102}

% ramsay1988monotone
\refj{Ramsay, J. O.}{1988}{Monotone regression splines in
  action}{Statistical Science}{3(4):425--441}

% schmidt1988positivity
\refj{Schmidt, J. W. and He{\ss}, W.}{1988}{Positivity of cubic
  polynomials on intervals and positive spline interpolation}{BIT
  Numerical Mathematics}{28(2):340--352}

% schumaker1983shape
\refj{Schumaker, Larry L.}{1983}{On shape preserving quadratic spline
  interpolation}{SIAM Journal on Numerical Analysis}{20(4):854--864}

% ulrich1994positivity
\refj{Ulrich, G. and Watson, L. T.}{1994}{Positivity conditions for
  quartic polynomials}{SIAM Journal on Scientific
  Computing}{15(3):528--544}

% wang2004rational
\refj{Wang, Q. and Tan, J.}{2004}{Rational quartic spline involving
  shape parameters}{Journal of Information and Computational
  Science}{1(1):127--130}

% yao2018unconditionally
\refj{Yao, J. and Nelson, K. E.}{2018}{An unconditionally monotone
  $\scriptstyle C^2$ quartic spline method with nonoscillation
  derivatives}{Advances in Pure Mathematics}{8(LLNL-JRNL-742107)}

}
%% -------------------------------------------------------------------

\vfill\eject\end