### 2022-03-11 [2020-08]_ACMTOMS_(MQSI)/paper/1-introduction.tex
Finishing revision draft, few more red highlights to add.

\heading{1. INTRODUCTION}

Many domains of science rely on smooth approximations to real-valued
functions over a closed interval. Piecewise polynomial functions
(splines) provide the smooth approximations for animation in graphics
[Herman et al.\ 2015; Quint 2003], aesthetic structural support in
architecture [Brennan 2020], efficient aerodynamic surfaces in
automotive and aerospace engineering [Brennan 2020], prolonged
effective operation of electric motors [Berglund et al.\ 2009], and
accurate nonparametric approximations in statistics [Knott
2012]. While polynomial interpolants and regressors apply broadly,
splines are often a good choice because they can approximate globally
complex functions while minimizing the local complexity of an
approximation.

It is often the case that the true underlying function or phenomenon
being modeled has known properties like convexity, positivity,
various levels of continuity, or monotonicity. Given a reasonable
amount of data, it quickly becomes difficult to achieve desirable
properties in a single polynomial function. In general, the
maintenance of function properties through interpolation/regression is
referred to as {\it shape preserving} [Fritsch and Carlson 1980;
Gregory 1985]. The specific properties the present algorithm will
preserve in approximations are monotonicity and $C^2$ continuity. In
addition to previously mentioned applications, these properties are
crucially important in statistics to the approximation of a cumulative
distribution function and subsequently the effective generation of
random numbers from a specified distribution [Ramsay 1988].  A spline
function with these properties could approximate a cumulative
distribution function to a high level of accuracy with relatively few
intervals. A twice continuously differentiable approximation to a
cumulative distribution function (CDF) would produce a corresponding
probability density function (PDF) that is continuously
differentiable, which is desirable.

The existing research in shape preserving interpolatory splines is
rich and filled with many approaches for many different
applications. In the context of this work, the unique quality of shape
preserving spline algorithms can be observed through: (1) the type of
spline (rational or Hermite), (2) the polynomial order and level of
continuity, and (3) the sharpness, sufficiency, and necessity of the
conditions used to establish shape preserving properties. While the
choice of spline representation is less important, achieving higher
levels of continuity is often more difficult than lower levels of
continuity. And while all referenced algorithms establish sufficient
conditions for shape preservation, it is less common and much more
difficult to provide sharp {\it necessary} conditions for
monotonicity. The following two paragraphs highlight a collection of
research related to the construction of sufficient conditions for
$C^1$ and $C^2$ shape preserving splines. From this foundation, the
value and utility of the code presented here can be more readily
appreciated.

DeVore [1977] proves that monotone spline interpolants have improved
accuracy over their nonmonotone counterparts when approximating
monotone functions, while Constantini [1986] proves that monotone
spline interpolants to data of fixed continuity and order exist in
general. This groundwork establishes the potential for monotone spline
interpolation. Schumaker [1983] and McAllister and Roulier [1981]
(independently) establish necessary and sufficient conditions for
monotone $C^1$ quadratic splines through the insertion of additional
knots and Fritsch [1982] establishes simplified sufficient conditions
for monotone $C^1$ cubic splines that require no knot insertion.
Gregory and Delbourgo [1983] prove necessary and sufficient conditions
for a closed form solution to monotone $C^1$ quadratic rational spline
interpolation with nonlinear boundary equations and an iterative
approach. Delbourgo and Gregory [1993] extend that same work to
achieve sufficient conditions for $C^2$ continuity with a cubic
rational spline given additional tension parameters defined by
users. Huynh [1993] similarly arrives at monotone $C^1$ cubic spline
interpolants and several necessary nonlinear boundary conditions on
monotonicity.

Continuing towards monotone $C^2$ spline interpolation, Fiorot and
Tabka [1991] prove a simple method for determining the existence of a
monotone $C^2$ cubic interpolating spline, but note that functions of
this kind do not always exist for arbitrary convex monotone data sets.
Pruess [1993] proposes sufficient conditions for a cubic $C^2$ shape
preserving spline method while acknowledging that quintic splines are
necessary for generally monotone $C^2$ spline interpolants that do not
insert new knots. Similarly, Manni [1997] constructs sufficient
conditions for a monotone $C^2$ cubic spline by adding two additional
knots per interval, then Cravery and Manni [2003] extend that
methodology to arrive at monotone $C^3$ interpolating splines by
progressively increasing tension through Bezier control
polygons. Constantini [1997] provides a method for monotone $C^2$
quintic spline interpolation based on sufficient and necessary
boundary value conditions, but the necessary conditions are not
sharp. Dougherty, Edelman, and Hyman [1989] construct monotone $C^2$
Hermite splines and in commentary recommend against the direct
optimization for smoothness parameters like global $L^2$, rather
promoting problem specific definitions of geometric niceness. Both
Wang and Tan [2004] as well as Yao and Nelson [2018] propose
sufficient conditions for $C^2$ quartic splines, while Piah and
Unsworth [2011] improve upon those sufficient conditions for $C^2$
quartic rational spline interpolation. Abbas, Majid, and Ali [2012]
formulate a set of sufficient conditions for monotone $C^2$ cubic
rational splines and similar work is extended to $C^2$ quartic
rationals (and splines with arbitrary smoothness) by Zhu and Han
[2015]. Ulrich and Watson [1994] arrived at sufficient and sharp
necessary conditions for monotone $C^2$ quintic splines, as also did
He{\ss} and Schmidt [1994] (independently). In addition, a statistical
method for bootstrapping the construction of an arbitrarily smooth
monotone fit exists [Leitenstorfer and Tutz 2006], but this is enabled
by sufficient conditions only.

The currently available peer reviewed and published software for
monotone piecewise polynomial interpolation is severely reduced in
comparison with the number of published approaches mentioned
above. This is partially indicative of how difficult it is to properly
handle the numerical conditions that arise when constructing and
evaluating precise shape preserving splines, as well as how difficult
it is to create code that correctly conforms with the theoretical
expectations. The sufficient $C^1$ cubic spline method PCHIP of
Fritsch and Carlson [1980] is available through the SciPy Python
package ($>1$ billion downloads), and the $C^1$ quadratic method of
Schumaker [1983] is available as an R package (300 thousand
downloads). The piecewise quadratic $C^1$ method of McAllister and
Roulier [1981] is available in FORTRAN (1000 downloads) and the
sufficient $C^2$ quintic method of Costantini [1997] BVSPIS in FORTRAN
as well (500 downloads). Based on publicly available download records,
it is assumed that the code by Fritsch [1982] for monotone $C^1$ cubic
spline interpolation is the predominant code for constructing monotone
interpolants at present.

The theory for sufficient and {\it sharp} necessary bounds for
monotone quintic interpolation has been provided [Ulrich and Watson
  1994; He{\ss} and Schmidt 1994] and that theory was recently
utilized in a proposed algorithm [Lux 2020] for monotone quintic
spline construction, however no published mathematical software exists
for the quintic case based on sharp monotonicity conditions. The
software presented here represents the first published software
package for producing $C^2$ shape preserving splines based on sharp
monotonicity conditions. This work improves upon the algorithm
presented by Lux et al. [2020] by refactoring computations for
improved numerical stability, estimating minimum magnitude second
derivatives at breakpoints with a quadratic facet model, and using a
binary search to reduce the magnitude of the modifications made to
initial derivative estimates when constructing a monotone spline
interpolant.

\heading{Overview}

This work provides a Fortran 2003 subroutine {\tt MQSI} based on the
sharp necessary and sufficient conditions in Ulrich and Watson [1994]
for the construction of monotone quintic spline interpolants of
monotone data. Precisely, the problem is, given a strictly increasing
sequence $X_1<X_2< \cdots <X_n$ of breakpoints with corresponding
monotone increasing function values $Y_1\le Y_2\le \cdots \le Y_n$,
find a $C^2$ monotone increasing quintic spline $Q(x)$ with the same
breakpoints satisfying $Q(X_i)=Y_i$ for $1\le i\le n$. ({\tt MQSI}
actually does something slightly more general, producing $Q(x)$ that
is monotone increasing (decreasing) wherever the data is monotone
increasing (decreasing).)

The remainder of this paper is structured as follows: Section 2 provides
the algorithms for constructing a $C^2$ monotone quintic spline interpolant
to monotone data, Section 3 outlines the method of spline representation
($B$-spline basis) and evaluation, Section 4 analyzes the complexity and
sensitivity of the algorithms in {\tt MQSI}, and Section 5 presents 
timing data and some graphs of constructed interpolants.