### 2022-03-19 [2020-08]_ACMTOMS_(MQSI)/paper/1-introduction.tex

\heading{1. INTRODUCTION}

Many domains of science rely on smooth approximations to real-valued
functions over a closed interval. Piecewise polynomial functions
(splines) provide the smooth approximations for animation in graphics
[Herman and Oftedal\ 2015; Quint 2003], aesthetic structural support in
architecture [Brennan 2020], efficient aerodynamic surfaces in
automotive and aerospace engineering [Brennan 2020], prolonged
effective operation of electric motors [Berglund et\ al.\ 2009], and
accurate nonparametric approximations in statistics [Knott
2012]. While polynomial interpolants and regressors apply broadly,
splines are often a good choice because they can approximate globally
complex functions while minimizing the local complexity of an
approximation.

It is often the case that the true underlying function or phenomenon
being modeled has known properties like convexity, positivity,
various levels of continuity, or monotonicity. Given a reasonable
amount of data, it quickly becomes difficult to achieve desirable
properties in a single polynomial function. In general, the
maintenance of function properties through interpolation/regression is
referred to as {\it shape preserving} [Fritsch and Carlson 1980;
Gregory 1985]. The specific properties the present algorithm will
preserve in approximations are monotonicity and $C^2$ continuity. 
\beginred
Notably this work does not consider convexity preservation, as Mulansky
and Neamtu [1991] prove that there exists strictly convex data sets which
do not allow convex interpolation with $C^2$ splines of fixed degree.
\endred
In addition to previously mentioned applications, \red{being $C^2$ and monotonicity preserving is}
crucially important in statistics to the approximation of a cumulative
distribution function and subsequently the effective generation of
random numbers from a specified distribution [Ramsay 1988].  A spline
function with these properties could approximate a cumulative
distribution function to a high level of accuracy with relatively few
intervals. A twice continuously differentiable approximation to a
cumulative distribution function (CDF) would produce a corresponding
probability density function (PDF) that is continuously
differentiable, which is desirable.

\beginred

The existing research in shape preserving interpolatory splines is
rich and filled with many approaches for many different
applications. In the context of this work, the unique quality of shape
preserving spline algorithms can be observed through: (1) the type of
spline (rational or Hermite), (2) the polynomial order and level of
continuity, and (3) the sharpness, sufficiency, and necessity of the
conditions used to establish shape preserving properties. While the
choice of spline representation is less important, achieving higher
levels of continuity is often more difficult than lower levels of
continuity. Furthermore, all referenced algorithms establish
sufficient conditions for shape preservation, but it is less common
and much more difficult to provide {\it sharp and necessary}
conditions for monotonicity. The following two paragraphs highlight a
collection of research related to the construction of $C^1$ and $C^2$
shape preserving splines. From this foundation, the value and utility
of the code presented here can be more readily appreciated.

DeVore [1977] proves that monotone spline interpolants have improved
accuracy over their nonmonotone counterparts when approximating
monotone functions, while Costantini [1986] proves that monotone
spline interpolants to data of fixed continuity and order exist in
general. This groundwork establishes the potential for monotone spline
interpolation. Schumaker [1983] and McAllister and Roulier [1981a]
(independently) establish necessary and sufficient conditions for
monotone $C^1$ quadratic splines through the insertion of additional
knots and Fritsch [1982] establishes simplified sufficient conditions
for monotone $C^1$ cubic splines that require no knot insertion.
Gregory and Delbourgo [1982] prove necessary and sufficient conditions
for a closed form solution to monotone $C^1$ quadratic rational spline
interpolation with nonlinear boundary equations and an iterative
approach. Delbourgo and Gregory [1983] extend that same work to
achieve sufficient conditions for $C^2$ continuity with a cubic
rational spline given additional tension parameters defined by
users. Huynh [1993] similarly arrives at monotone $C^1$ cubic spline
interpolants and several necessary nonlinear boundary conditions on
monotonicity.

Continuing towards monotone $C^2$ spline interpolation, Fiorot and
Tabka [1991] prove a simple method for determining the existence of a
monotone $C^2$ cubic interpolating spline, but note that functions of
this kind do not always exist for arbitrary convex monotone data sets.
Pruess [1993] proposes sufficient conditions for a cubic $C^2$ shape
preserving spline method while acknowledging that quintic splines are
generally necessary for monotone $C^2$ spline interpolants that do not
insert new knots. Similarly, Manni and Sablonni\`ere [1997] construct
sufficient conditions for a monotone $C^2$ cubic spline by adding two
additional knots per interval, then Cravero and Manni [2003] extend
that methodology to arrive at monotone $C^3$ interpolating splines by
progressively increasing tension through Bezier control
polygons. Costantini [1997] provides a method for monotone $C^2$
quintic spline interpolation based on sufficient and necessary
boundary value conditions, but the necessary conditions are not
sharp. Dougherty, et\ al.\ [1989] construct monotone $C^2$ Hermite
splines and in commentary recommend against the direct optimization
for smoothness parameters like global $L^2$, rather promoting problem
specific definitions of geometric niceness. Both Wang and Tan [2004]
as well as Yao and Nelson [2018] propose sufficient conditions for
$C^2$ quartic splines, while Piah and Unsworth [2011] improve upon
those sufficient conditions for $C^2$ quartic rational spline
interpolation. Abbas, et\ al.\ [2012] formulate a set of sufficient
conditions for monotone $C^2$ cubic rational splines and similar work
is extended to $C^2$ quartic rationals (and splines with arbitrary
smoothness) by Zhu and Han [2015a; 2015b]. Ulrich and Watson [1990]
arrive at sharp necessary and sufficient conditions for monotone $C^2$
quintic splines, while He{\ss} and Schmidt [1994] produce similar
sufficient conditions for $C^2$ quintic splines. In addition, a method
for constructing an arbitrarily smooth monotone fit exists
[Leitenstorfer and Tutz 2006] as well as a method for arbitrarily
smooth shape preservation [Han 2018], but these methods are enabled by
sufficient conditions only.

The currently available peer reviewed and published software for
monotone piecewise polynomial interpolation is severely limited in
comparison with the number of published approaches mentioned
above. This is partially indicative of how difficult it is to properly
handle the numerical conditions that arise when constructing and
evaluating precise shape preserving splines, as well as how difficult
it is to create code that correctly conforms with the theoretical
expectations. The sufficient $C^1$ cubic spline method PCHIP of
Fritsch and Carlson [1980] is available through the SciPy Python
package ($>$1 billion downloads), and the $C^1$ piecewise quadratic
method of Schumaker [1983] is available as an R package (300,000
downloads). The $C^1$ piecewise quadratic method of McAllister and
Roulier [1981a;\ 1981b] is available in FORTRAN (1000 downloads) and
the sufficient $C^2$ piecewise quintic method of Costantini [1997a;
  1997b] BVSPIS in FORTRAN as well (500 downloads). Based on publicly
available download records, it is assumed that the code by Fritsch
[1982] for monotone $C^1$ cubic spline interpolation is the
predominant code for constructing monotone interpolants at present.

\endred

The theory for \red{{\it sharp}} necessary and sufficient bounds for
monotone $C^2$ quintic interpolation has been provided [Ulrich and
  Watson 1990; 1994] and that theory was recently utilized in a
proposed algorithm [Lux et al. 2020; Lux 2020] for monotone $C^2$
quintic spline construction, however no published mathematical
software exists for the quintic case \red{based on sharp monotonicity
conditions}. The software presented here represents the first published
software package for producing $C^2$ shape preserving splines
\red{based on sharp monotonicity conditions}. This work improves upon
the algorithm presented by Lux et\ al.\ [2020] by refactoring
computations for improved numerical stability, estimating minimum
magnitude second derivatives at breakpoints with a quadratic facet
model, and using a binary search to reduce the magnitude of the
modifications made to initial derivative estimates when constructing a
monotone spline interpolant.

\beginred

\heading{Overview}

\endred

This work provides a Fortran 2003 subroutine {\tt MQSI} based on the
sharp necessary and sufficient conditions in Ulrich and Watson [1994]
for the construction of monotone $C^2$ quintic spline interpolants of
monotone data. Precisely, the problem is, given a strictly increasing
sequence $X_1<X_2< \cdots <X_n$ of breakpoints with corresponding
monotone increasing function values $Y_1\le Y_2\le \cdots \le Y_n$,
find a $C^2$ monotone increasing quintic spline $Q(x)$ with the same
breakpoints satisfying $Q(X_i)=Y_i$ for $1\le i\le n$. ({\tt MQSI}
actually does something slightly more general, producing $Q(x)$ that
is monotone increasing (decreasing) wherever the data is monotone
increasing (decreasing).)

The remainder of this paper is structured as follows: Section 2
provides the algorithms for constructing a monotone $C^2$ quintic
spline interpolant to monotone data, Section 3 outlines the method of
spline representation ($B$-spline basis) and evaluation, Section 4
analyzes the complexity and sensitivity of the algorithms in {\tt
  MQSI}, and Section 5 presents timing data and some graphs of
constructed interpolants \red{as well as a visual comparison with
  existing monotone spline packages}.