### 2022-10-26 [2020-08]_ACMTOMS_(MQSI)/paper/LaTeX/3-spline-rep.tex
Adding LaTeX version of paper.

\section{SPLINE REPRESENTATION}
The monotone quintic spline interpolant $Q(x)$ is represented in terms
of a B-spline basis. The routine {\tt FIT\_SPLINE} in this package
computes the B-spline coefficients $\alpha_i$ of $Q(x)=\sum_{i=1}^{3n}
\alpha_i B_{i,6,t}(x)$ to match the piecewise quintic polynomial
values and (first two) derivatives at the breakpoints $X_i$, where the
spline order is six and the knot sequence $t$ has the breakpoint
multiplicities (6, 3, $\ldots$, 3, 6).  The routine {\tt EVAL\_SPLINE}
evaluates a spline represented in terms of a B-spline basis. A Fortran
2003 implementation {\tt EVAL\_BSPLINE} of the B-spline recurrence
relation evaluation code by C. de Boor [1978] for the value,
derivatives, and integral of a B-spline is also provided.